Foundations for Inference

### 2018-10-19 DATA606/Foundations for Inference.Rmd
Additional work on homework

(b) Suppose you read online that children first count to 10 successfully when they are 32 months old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence that the average age at which gifted children first count to 10 successfully is less than the general average of 32 months. Use a significance level of 0.10.

(c) Interpret the p-value in context of the hypothesis test and the data.

(d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.

(e) Do your results from the hypothesis test and the confidence interval agree? Explain.

@@

@@

@@

(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Centered around 9,000 hours.**

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours? **It would be roughly 0.067**

@@

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Yes, but not using the above method.**

>>>>

(b) Suppose you read online that children first count to 10 successfully when they are 32 months old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence that the average age at which gifted children first count to 10 successfully is less than the general average of 32 months. Use a significance level of 0.10. **$H_0$ is that x_bar >= 32 and $H_a$ is x_bar < 32**

```{r}
alpha <- 0.10
n <- 36
x_bar <- 30.69
s <- 4.31
mu <- 32

se_x_bar <- s / sqrt(n)

test_stat <- (x_bar - mu)/se_x_bar

t_alpha <- qt(1-alpha, df = n-1) * -1

p_value <- pt(test_stat, df=n-1)

if(test_stat < t_alpha){
  message(paste(test_stat, "is less than", t_alpha, "so reject the null hypothesis (p-value:", p_value, ")"))
} else {
  message(paste(test_stat, "is greater than", t_alpha, "so accept the null hypothesis (p-value:", p_value, ")"))
}
```

(c) Interpret the p-value in context of the hypothesis test and the data.  **There is a 4% probability that the difference between the two means would be equal to or more extreme than its observed value.**

(d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.
```{r}
error <- qt(1-alpha, df=n-1) * se_x_bar
ci_low <- x_bar - error
ci_high <- x_bar + error
```
**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**

(e) Do your results from the hypothesis test and the confidence interval agree? Explain. **Yes since 32 is outside of the confidence interval you would reject the null hypotheses.  it is the same conclusion you reach with the hypothesis test.**

@@

```{r}
alpha <- 0.10
n <- 36
x_bar <- 118.2
s <- 6.5
mu <- 100

se_x_bar <- s / sqrt(n)

test_stat <- (x_bar - mu)/se_x_bar

t_alpha <- qt(1-alpha, df = n-1) * -1

p_value <- pt(test_stat, df=n-1)

if(test_stat < t_alpha){
  message(paste(test_stat, "is less than", t_alpha, "so reject the null hypothesis (p-value:", p_value, ")"))
} else {
  message(paste(test_stat, "is greater than", t_alpha, "so accept the null hypothesis (p-value:", p_value, ")"))
}
```

@@

```{r}
error <- qt(1-alpha, df=n-1) * se_x_bar
ci_low <- x_bar - error
ci_high <- x_bar + error
```

**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**

@@

(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Normal and centered around 9,000 hours.**

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?

@@

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Not using the above method.**
